### 2017-02-13 Documents/bootcamp_practice_pyfund2.tex
add practice_pyfund2

\documentclass[11pt]{exam}

\input{../LaTeX/preamble.tex}

%\printanswers

% document starts here
\begin{document}
%\parskip=\bigskipamount
\parindent=0.0in
\thispagestyle{empty}
%\input{../LaTeX/header.tex}

\bigskip\bigskip
\centerline{\Large \bf Data Bootcamp:  Python Fundamentals 1 Practice}
\centerline{Revised: \today}

\medskip

\vspace{2cm}

\begin{questions}
\item Construct a dictionary whose keys are the integers $1$, $2$ and $3$ and whose values are the same numbers as words: one, two, three. How would you get the word associated with the key 2?

\item What does the built-in function \texttt{zip()} do? Start with the lists
\begin{verbatim}
l1 = [1, 2, 3]
l2 = ['one', 'two', 'three']
\end{verbatim} 
\begin{itemize}
\item Construct the list of tuples \texttt{[(1, 'one'), (2, 'two'), (3, 'three')]}.
\item Construct a dictionary whose keys are the integers $1$, $2$ and $3$ and whose values are the same numbers as words: one, two, three.  
\end{itemize}

\item Take the variables \texttt{name1} and \texttt{name2}, both of them strings. Write a program using \texttt{if ... else} that prints the name that comes first in alphabetical order. Test your program with 
\begin{verbatim}
name1 = 'Dave'
name2 = 'Glenn'
\end{verbatim}
Could you write a function for this task?

\item Take the list \texttt{stuff = ['cat', 3.7, 5, 'dog']}. 
\begin{itemize}
\item Write a program that tells us the type of each elements of \texttt{stuff} (value and type)
\item Write a program that prints only the elements of \texttt{stuff} that are strings. 
\end{itemize}

\item Write a program that given two integer values \texttt{a} and \texttt{b} returns \texttt{True} if either one is $6$, or if their sum or difference is $6$. \emph{Hint:} you might want to use the built-in function \texttt{abs()}. Test your program with
\begin{itemize}
\item \texttt{a=6}, \texttt{b=4} (should get \texttt{True})
\item \texttt{a=4}, \texttt{b=5} (should get \texttt{False})
\item \texttt{a=1}, \texttt{b=5} (should get \texttt{True})
\end{itemize}
Could you write a function for this task?

\item Write a program that given two integers \texttt{a} and \texttt{b} returns their sum. However, sums in the range $[10, 19]$ are forbidden, so in that case just return $20$. Test your program with
\begin{itemize}
\item \texttt{a=3}, \texttt{b=4} (should get $7$)
\item \texttt{a=9}, \texttt{b=4} (should get $20$)
\item \texttt{a=10}, \texttt{b=11} (should get $21$)
\end{itemize}
Could you write a function for this task?

\item Take the list of bond yields \texttt{y = [.01, .02, .03]} for maturities of one, two and three years. Write a program that multiplies all elements with 100, i.e. returns the list 
\begin{verbatim}
y_perc = [1.0, 2.0, 3.0]
\end{verbatim}
Can you do it with list comprehension? And with a function?

\item Using only core \texttt{Python} functionality define a list that contains 
\begin{itemize}
\item all integers from 0 to \texttt{n} (test it with the example: \texttt{n=5})
\item all integers from \texttt{n1} to \texttt{n2} (test it with the example: \texttt{n1=4} and \texttt{n2=8}) 
\end{itemize}
Could you define a function that can do both? (\emph{Hint:} use default values)

\begin{center}
\noindent\rule{10cm}{0.4pt}
\end{center}

\item Suppose that you have two strings
\begin{verbatim}
s1 = "abc def ghi"
s2 = "def ghi abc"
\end{verbatim}
\begin{itemize}
\item How would you check if the two strings contain the same characters? (irrespective of their order)
\item How would you check if the two strings contain the same words?
\end{itemize}

\item Take the list \texttt{x = [1, 4, 5, 6, 4, 2, 7, 12, 4, 23, 4]}. Using only loops and/or conditionals create a list with all the indices \texttt{i} for which 
\texttt{x[i] == 4}.

\end{questions}
%\input{../LaTeX/footer.tex}

\end{document}