### 2017-02-13 Documents/bootcamp_practice_pyfund1.tex

\documentclass[11pt]{exam}

\input{../LaTeX/preamble.tex}

%\printanswers

% document starts here
\begin{document}
%\parskip=\bigskipamount
\parindent=0.0in
\thispagestyle{empty}
%\input{../LaTeX/header.tex}

\bigskip\bigskip
\centerline{\Large \bf Data Bootcamp:  Python Fundamentals 1 Practice}
\centerline{Revised: \today}

\medskip
%{\it Answer each of the questions below.
%Include any code you write.
%What you hand in can be code or handwritten or something else,
%but it must be readable and professional.
%Hardcopy only.}
\vspace{2cm}

\begin{questions}
\item Fix this code:
\begin{verbatim}
bad_string = 'Sarah's code'
\end{verbatim}

\item How would you convert the integer i = 1234 to the list 
\begin{verbatim}
l = ['1', '2', '3', '4']
\end{verbatim}

\item Suppose \textbf{year} is a string containing the year of a particular piece of data; for example, 
\begin{verbatim}
year = '2016'
\end{verbatim}
How would we construct a \emph{string} for the following year?

\item Suppose we have two variables, $x$ and $y$. How would you switch their values, so that $x$ takes on $y$'s value and $y$ takes on $x$'s?

\item Set 
\begin{verbatim}
name = 'Jones'
\end{verbatim}
Use (a) tab completion to find a method that coverts \textbf{name} to upper case (capital) letters and (b) the Object inspector to find out how to use that method.

\item Set 
\begin{verbatim}
name = 'Ulysses'
\end{verbatim}
Use (a) tab completion to find a method that counts the number of appearances of the letter $s$ and (b) the Object inspector to find out how to use that method.

\item Set
\begin{verbatim}
long_string = 'salamandroid'
\end{verbatim}
Find a method to replace the $a$'s with asterisks.

\item Create the list 
\begin{verbatim}
list = [1, 2, 3, 4, 5]
\end{verbatim}
and using list methods
\begin{itemize}
\item reverse the elements of $l$ and print it
\item remove the number $4$ from $l$ and print it 
\item append the float $1.5$ to $l$ and print it
\item sort the elements of $l$ and print it
\end{itemize}

\begin{center}
\noindent\rule{10cm}{0.4pt}
\end{center}

\pagebreak

\item You are given the number (as a string) in the following form
\begin{verbatim}
number = "32,054.23"
\end{verbatim}
How would you convert this into a valid float? You should get $32054.23$.

\item You are given a name of the form 
\begin{verbatim}
firstname_lastname = 'john_doe'
\end{verbatim}
Save Firstname and Lastname variables with capitalized values using string methods.

\item You are given a list of the form: $l = [0, 1, 2, 4, 5]$. Use methods for lists to produce a corrected list which contains the number $3$ between $2$ and $4$:

\begin{verbatim}
l = [0, 1, 2, 3, 4, 5]
\end{verbatim}

\item You are given a list of the form: $l = [\dots, 1, 2, 4, 5, \dots]$ (with an ``unkown'' number of first and last elements). Use methods for lists to produce a corrected list which contains the number $3$ between $2$ and $4$:

\begin{verbatim}
l = [..., 1, 2, 3, 4, 5, ...]
\end{verbatim}

\item You are given strings of web domains of the form: $s = $'www.example.com'. You want to strip the \textbf{'www.'} and the \textbf{'.com'}. Use string methods to achieve this.

\item Write a Python program to get the part of the string 
\begin{verbatim}
link = 'https://play.spotify.com/collection/albums'
\end{verbatim}
before the last '/' character. You should get \begin{verbatim}
'https://play.spotify.com/collection'
\end{verbatim}

\item \textbf{Hard!} Write a program to swap comma and dot in a string. In particular turn the string
\begin{verbatim}
dot_comma = "32.054,23"
\end{verbatim}
into
\begin{verbatim}
comma_dot = "32,054.23"
\end{verbatim}
\emph{Hint:} look up the string methods \textbf{maketrans} and \textbf{translate}.
\end{questions}

%\input{../LaTeX/footer.tex}

\end{document}