### 2017-09-04 Documents/bootcamp_3ideas.tex
Outline and ideas change

%%%%%%%%%%%%%%%%%%%%%% Stock Preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{../LaTeX/undergrad_preamble.tex}

%%%%%%%%%%%%%%%%%%%%%%Headers and Footers%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{headandfoot}
\runningheadrule
\firstpageheadrule
\firstpageheader{\includegraphics[width=0.25\textwidth]{../Figures/stern_black1.pdf}}{}{Data Bootcamp: Three Project Ideas}
\runningheader{}{}{}
\runningfooter{}{}{}

\begin{document}
\bigskip
\centerline{\Large \bf Data Bootcamp: Three Project Ideas}
\medskip
\centerline{Revised: \today}

\bigskip
{\it Write down three project ideas, a sentence or two each. We'll use them for class discussion. 
Save for later, they're a head start on coming up with a project idea.}

%\input{../LaTeX/footer.tex}

\end{document}